---
title: "Unknown_analysis_single_cell"
author: "Filippo Mirolo"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---
## Dependencies

```{r}
set.seed(62)
library(tidyverse)
library(Seurat)
library(patchwork)
library(celldex)
library(SingleR)
library(SummarizedExperiment)
```

## Data exploration

```{r}
setwd("C:\\Users\\filoa\\OneDrive\\Desktop\\Single_cell")
data <- readRDS("Single_cell_project/data/EOB.rds")
dim(data)
colnames(data)[1:15]
```
  Starting creating a `Seurat` object, with `CreateSeuratObject` needs the count matrix and
further parameters to filter and cut out some genes or cell which are useless. I'm passing to it
the min number of genes (features) and (cells) to be kept.
```{r}
dat <- CreateSeuratObject(data, min.cells = 8, min.features = 70)
```
  Now i continue with the exploration, I will looking for number of unique genes between cells 
(these are quite no informative for all the analysis because bring noise and can't be compared), 
doublets, total number of molecules (will be useful later for normalization). High variability of 
the total number of molecules make the analysis very unstable, in fact can introduce errors, also 
quite important in the normalization process. Lastly the number of mitochondrial molecules are an 
index of damaged cells, which can have altered pattern of transcription. ERCC external RNA control consortium.
```{r}
dat[["percent_mt"]] <- PercentageFeatureSet(dat, pattern = "^ERCC-")
head(dat@meta.data, 5)
sum(dat@meta.data$percent_mt)
```

```{r}
d <- median(dat$nCount_RNA)
```

```{r}
p1<- VlnPlot(dat, features = c("nFeature_RNA", "nCount_RNA", "percent_mt"),
             ncol = 3, pt.size = 0.01)
p1
  
plot1 <- FeatureScatter(dat, feature1 = "nCount_RNA", feature2 = "percent_mt")
plot2 <- FeatureScatter(dat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
  
plot1 + plot2

```
  Removing those who have too mach mitochondrial RNA and those who have not a good amount of features.
```{r}
dat <- subset(dat, subset = percent_mt < 4 &
              nFeature_RNA > 500 &
                nCount_RNA < 2*d)
```

```{r}
p1<- VlnPlot(dat, features = c("nFeature_RNA", "nCount_RNA", "percent_mt"),
             ncol = 3, pt.size = 0.01)
p1
  
plot1 <- FeatureScatter(dat, feature1 = "nCount_RNA", feature2 = "percent_mt")
plot2 <- FeatureScatter(dat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
  
plot1 + plot2

```
Scaling apply a z-score transformation (normalization) across cells. This is meant to give the same weight to all the cells, accessible thanks to @scale.data.

  I really don't understand which are the assumptions behind the normalization?
Why we can scale the data?

```{r}
dat <- NormalizeData(dat)   # normalization
dat <- FindVariableFeatures(dat)  # feature selection
dat <- ScaleData(dat)
```
  Cell cycle detection is really important in this kind of analysis, where there is lot of 
variability. In fact the same cell types can have different transcription pattern if they are in different phrases.
```{r}
s_genes <- cc.genes$s.genes    # S phrase reference
g2m_genes <- cc.genes$g2m.genes   # G2M phrase reference

#cell_cycle <- CellCycleScoring(dat, s.features = s_genes, g2m.features = g2m_genes, set.ident = F)
#cell_cycle <- RunPCA(cell_cycle, features = c(s_genes, g2m_genes))
#DimPlot(cell_cycle)
```

```{r}
dat <- RunPCA(dat, features = VariableFeatures(object = dat), verbose = F)
ElbowPlot(dat)
```

```{r}
dat <- RunUMAP(dat, reduction = "pca", dims = 1:20)
DimPlot(dat)
dat <- FindNeighbors(dat, reduction = "pca", dims = 1:20)
dat <- FindClusters(dat)
DimPlot(dat)
```



```{r}
ref <- MouseRNAseqData()
length(ref$label.main)
ref$label.main[1:30]
```

```{r}
pred <- SingleR(test=dat[["RNA"]]@counts,
                ref=ref, labels=ref$label.main, fine.tune = TRUE, prune = TRUE)
length(pred)
```

```{r}
table(pred$labels)
table(pred$pruned.labels)
table(is.na(pred$pruned.labels))
```

```{r}
plotScoreHeatmap(pred)
```

```{r}
identical(colnames(dat),rownames(pred))
dat@meta.data$SingleR_label <- pred$pruned.labels

DimPlot(dat, group.by = "SingleR_label")
DimPlot(dat)
```
